### 2023-04-20 paper.Rmd
editing

To optimize overall between-group similarity---and hence to overcome the limitations of k-means anticlustering---this paper introduces k-plus anticlustering. K-plus is an extension of the k-means objective that quantifies between-group similarity as discrepancy with regard to several distribution characteristics instead of only the mean. Specifically, k-plus offers the possibility to minimize differences with regard to the variance and higher order moments such as skewness and kurtosis. After formally introducing the k-plus objective, a simulation study shows that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming traditional anticlustering criteria. Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust` [@papenberg2020; @R-anticlust], an extension to the popular statistical programming language R [@R-base].

@@

In the following, I present a variation of the $\mathit{SSE}$ that can be used to minimize differences in variance between groups instead of only mean values. To motivate the new criterion, we first consider a set of one-dimensional observations. The sample variance of a set of $N$ observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:

@@

  # 3. SSE(KPLUS)
  # Use standardization to compute k-plus criterion!
  all_objectives_kPlus <- sapply(
    partitions,
    FUN = var_objective,
    features = cbind(features, squared_distances_from_mean)
  )

@@

Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict with each other [@ferligoj1992direct]. As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown whether maximizing it can yield a partitioning where both constituting criteria (mean and variance) are similar between groups. A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible. I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups. Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions. The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$. Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (represented by the triangular symbol) is not well-suited to optimize $\mathit{SSE}$ at the same time. Thus, when only considering one of the objectives, the mathematically optimal partitioning fails to establish groups that are similar both with regard to their means and variances. The diamond symbol represents the partition that maximizes the unweighted sum of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ (i.e., $w_1$ and $w_2$ were set to 1 when computing $\mathit{SSE_{kplus}}$), which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$. Thus, even using a very simple unweighted sum approach, the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ can be fulfilled to an astonishing degree, indicating that these objectives tend to align and are not in conflict with each other. Moreover, the scatterplot in Figure 1 shows that there is a high concentration of partitions in the upper-right quadrant where the two objectives are in high accordance.

@@

To formalize the computation of the k-plus criterion for an arbitraty number of distribution moments, some new notation has to be provided. First, we define the $j$'th column ($j = 1, \ldots, P$) of the original data matrix $\mathbf{X}_{N \times P}$ as $x_{.j} =  (x_{1j}, \ldots, x_{Nj})$ where the column mean is $\overline{\mathbf{x}}_j$. Let $T$ be the number of distribution moments that should be equalized between $K$ groups. For each moment $t$ ($t = 1, \ldots, T$) and for each variable $x_{.j}$, a new *t-moment variable* $\hat{x}_{t.j} = (\hat{x}_{t1j}, \ldots, \hat{x}_{tNj})$ is defined where

@@

Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{x}_{tj} ^{(k)}$ is mean for the $k$'th group on the $j$'th $t$-moment variable. Maximizing $\mathit{SSE_{kplusGeneralized}}$ can be used to equalize the first $T$ distribution moments between $K$ groups.

@@

When defining the row vectors of $\mathbf{X'}_{N \times \mathit{TP}}$ with the notation $x'_{i.} = (x'_{i1}, \ldots, x'_{i(TP)})$ for each $i = 1, \ldots, N$, the criterion $\mathit{SSE_{kplusGeneralized}}$ can be computed as

@@

In general, the weighting scheme for k-plus anticlustering can become highly complex as it might encompass an arbitrary number of distribution moments, and hence, an arbitrary number of weights. While an exhaustive evaluation of the influence of weights is beyond the scope of the current paper, a small scale simulation study was conducted to investigate the influence of weights for a selected number of moments. The target was the optimization of $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$, i.e. minimizing differences with regard to means, variances, skew and kurtosis. Two unweighted approaches were implemented by defining $\mathbf{w} = (1, 1, 1, 1)$. The unweighted standardization approach implemented a $z$-standardization of all columns---the original as well as the augmented variables---of the input data $\mathbf{X'}_{N \times \mathit{TP}}$ before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized. In the second case, the unstandardized matrix $\mathbf{X'}_{N \times \mathit{TP}}$ was used. Four weighted approaches were implemented by giving a tenfold increased weight to one of the four moments, respectively. That is, to compute $\mathit{SSE_{kplusGeneralized}}$ favoring similarity in means during the optimization process, I set $\mathbf{w} = (10, 1, 1, 1)$; the weights favoring similarity in variances, skewness and kurtosis, were defined as $\mathbf{w} = (1, 10, 1, 1)$, $\mathbf{w} = (1, 1, 10, 1)$, and $\mathbf{w} = (1, 1, 1, 10)$, repectively. The weighted approaches did not use a standardization of the input data.

To compare the different weighting schemes, `r length(results)` data sets from were generated from a normal distribution ($M = 0$, $\mathit{SD} = 1$). The number of variables $P$ was determined randomly for each data set and varied between 1 and 5; the number of anticlusters $K$ was also determined randomly for each data set and varied between 2 and 4; the sample size $N$ randomly varied between 20 and 100 with the restriction that $N$ had to be divisible by $K$. For each data set, $\mathit{SSE_{kplusGeneralized}}$ was optimized using the local maximum algorithm, once for each of the six competing weighting schemes.

```{r}
final_tab_sim1 <- round(Reduce("+", results) / length(results) * 100, 2)
rownames(final_tab_sim1) <- c(
  "$\\mathbf{w} = (1, 1, 1, 1)$ / unstandardized",
  "$\\mathbf{w} = (1, 1, 1, 1)$ / standardized",
  "$\\mathbf{w} = (10, 1, 1, 1)$",
  "$\\mathbf{w} = (1, 10, 1, 1)$",
  "$\\mathbf{w} = (1, 1, 10, 1)$",
  "$\\mathbf{w} = (1, 1, 1, 10)$"
)
colnames(final_tab_sim1) <- c(
  "$\\mathit{SSE}$",
  "$\\mathit{SSE_{Var}}$", 
  "$\\mathit{SSE_{Skew}}$",
  "$\\mathit{SSE_{Kurtosis}}$",
  "Average"
)
note <- paste0("The results of Simulation 1 averaged across all ", length(results), " data sets. Cells indicate how well the maximization of the weighted $\\mathit{SSE_{kplusGeneralized}}$ fulfilled each of the four constituting criteria $\\mathit{SSE}$, $\\mathit{SSE_{Var}}$, $\\mathit{SSE_{Skew}}$, and $\\mathit{SSE_{Kurtosis}}$. Cell values represent relative performance, i.e. they were computed as the $\\mathit{SSE}$ that was attained by the given weighting scheme, divided by the best $\\mathit{SSE}$ that was obtained by any weighting scheme in a given simulation run.")

apa_table(
  final_tab_sim1, 
  escape = FALSE, 
  caption = "Results of Simulation study 1",
  note = note,
  stub_indents = list("Unweighted approaches" = 1:2,
                      "Weighted approaches" = 3:6)
)

@@

Table 1 illustrates how well each of the six weighting schemes fulfilled each of the four constititung criteria, averaged across the `r length(results)` simulation runs. It is shown that for the weighted approaches, the criterion that received the highest weight was fulfilled to the highest degree, as should be expected. Interestingly, the unweighted standardization approach had the best average performance across the four criteria. While the degree of advantage over the other weighting schemes by itself does not seem to be remarkable, this pattern was observed in an astonishing `r mean(unstd_best) * 100`% of all simulation runs (the level of chance would have been `r 1/6 * 100`%, `r format_p(sig_test$p.value)`). Among the competing weighting schemes, the unweighted standardization approach therefore yielded the most promising results. It also yielded satisfying results with regard to maximizing the standard $\mathit{SSE}$, comparable to the weighting scheme explicitly favoring $\mathit{SSE}$ (**TODO**: CHECK IF THESE RESULTS HOLD WHEN ALL DATA IS IN!). Hence, the unweighted standardization approach seemed to implicitly give high priority to minimizing differences in means. This is a pleasant result because, as discussed earlier, for the higher order moment $\mathit{SSE}$s to be valid, the group means have to be close to the overall mean, which is best accomplished if the standard $\mathit{SSE}$ has high priority.

@@

A simulation study and examples on real norming data showed that the k-plus criterion is well-suited to maximize between-group similarity with regard to multiple criteria. Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, k-plus anticlustering should be considered as the default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be challenging in small data sets. In practice, users can try out different objectives and see which specification fulfills their needs. A reasonable approach is to start "greedy", e.g. by specifying that multiple distribution moments should equalized. If the results of such a greedy setup are not satisfying, users can reduce their requirements by opting for fewer criteria (e.g., only the mean and variance). Using the open source R package `anticlust`, the (repeated) application of k-plus anticlustering is easily accessible and free.

@@

The present paper focused on the new k-plus criterion for anticlustering, but only superficially considered the multicriterion nature of this criterion. A simple unweighted sum approach was used to optimize the k-plus objective, while more sophisticated algorithmic procedures exist to tackle multicriterion optimization problems [e.g., @brusco2009cross]. The sole focus on the objective function followed two reasonings. First, the applications showed that the local maximum search, maximizing the unweighted k-plus sum, yielded satisfying results that probably suffice in most applications. It is possible that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfill multiple criteria to a satisfactory degree (as suggested by Figure 1). Still, more sophisticated approaches may yield theoretically improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives. @brusco2019 gave an example of how direct algorithms can be employed to optimize multiple criteria in anticlustering applications aiming for high within-group heterogeneity. Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment. A second reason for focusing on the k-plus objective per se---independently from the algorithm that is used to optimize it---was to obtain a fair comparison to existing anticlustering objectives. In the simulation study, differences in performance could not have been interpreted if the objective function as well as the algorithm used to optimize it had differed between anticlustering methods. It was of particular interest to compare the k-plus objective with the popular diversity objective. @papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should be preferred because k-means does not equalize the spread of the distribution between groups. In contrast, maximizing the diversity leads to an appropriate balance between location and spread. Using the k-plus extension, however, the k-means criterion has regained attractivity: When opting for similarity in means and variances---and for increasing $N$, skewness and kurtosis as well---, k-plus anticlustering clearly outperformed diversity anticlustering.

It should be noted that because the covariance is defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariance, a possibility that was not investigated in the current paper. The covariance between two variables $x = (x_1, \ldots, x_N)$ and $y = (y_1, \ldots, y_N)$ is defined as $\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$. Thus, to maximize between-group similarity with regard to covariance structure, it is possible to augment the data set with an additional variable for each pair of variables $x$ and $y$, computed as $(x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$. Future research should examine how the k-plus approach can be adapted to including covariances, including a discussion of how to select weights for the possibly high number of additional variables. The unweighted standardization approach presented here may serve as a starting point for investigation.

>>>>

To optimize overall between-group similarity---and hence to overcome the limitations of k-means anticlustering---this paper introduces k-plus anticlustering. K-plus is an extension of the k-means objective that quantifies between-group similarity as discrepancy with regard to several distribution characteristics instead of only the mean. Specifically, k-plus offers the possibility to minimize differences with regard to the variance and higher order moments such as skewness and kurtosis. After formally introducing the k-plus objective, two simulation studies show that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming traditional anticlustering criteria. Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust` [@papenberg2020; @R-anticlust], an extension to the popular statistical programming language R [@R-base].

@@

In the following, I present a variation of the $\mathit{SSE}$ that can be used to minimize differences in variance between groups instead of only means. To motivate the new criterion, we first consider a set of one-dimensional observations. The sample variance of a set of $N$ observations $x = (x_1, \ldots, x_N)$ is computed as the mean of the squared distances between observations and their mean $\overline{\mathbf{x}}$:

@@

  # 3. SSE(KPLUS)
  all_objectives_kPlus <- sapply(
    partitions,
    FUN = var_objective,
    features = cbind(features, squared_distances_from_mean)
  )

@@

Research on multiobjective optimization indicates that solutions oftentimes fail to satisfactorily address several criteria if the underlying objectives are in conflict with each other [@ferligoj1992direct]. As $\mathit{SSE_{kplus}}$ is a novel criterion, it is so far unknown whether maximizing it can yield a partitioning where both constituting criteria are similar between groups. A motivating example illustrates that aiming to simultaneously optimize similarity of means and variances may be feasible. I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups. Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions. The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$. Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (represented by the triangular symbol) is not well-suited to optimize $\mathit{SSE}$ at the same time. Thus, when only considering one of the objectives, the mathematically optimal partitioning fails to establish groups that are similar both with regard to their means and variances. The diamond symbol represents the partition that maximizes the unweighted sum of $\mathit{SSE_{Var}}$ and $\mathit{SSE}$ (i.e., $w_1$ = $w_2$ = 1 in the computation of $\mathit{SSE_{kplus}}$), which yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$. Thus, even using a very simple unweighted sum approach, the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ can be fulfilled to an astonishing degree. This result indicates that the objectives $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ are not in conflict with each other. Moreover, in Figure 1 there is a high concentration of partitions in the upper-right corner where the two objectives are in high agreement.

@@

To formalize the computation of the k-plus criterion for an arbitrary number of distribution moments, some new notation has to be provided. First, we define the $j$'th column ($j = 1, \ldots, P$) of the original data matrix $\mathbf{X}_{N \times P}$ as $x_{.j} =  (x_{1j}, \ldots, x_{Nj})$ where the column mean is $\overline{\mathbf{x}}_j$. Let $T$ be the number of distribution moments that should be equalized between $K$ groups. For each moment $t$ ($t = 1, \ldots, T$) and for each variable $x_{.j}$, a new *t-moment variable* $\hat{x}_{t.j} = (\hat{x}_{t1j}, \ldots, \hat{x}_{tNj})$ is defined where

@@

Here, $w_t$ is the weight associated with the $t$'th moment and $\overline{x}_{tj} ^{(k)}$ is mean for the $k$'th group on the $j$'th $t$-moment variable. Maximizing $\mathit{SSE_{kplusGeneralized}}$ can be used to minimize differences with regard to the first $T$ distribution moments between $K$ groups.

@@

When defining the row vectors of $\mathbf{X'}_{N \times \mathit{TP}}$ with the notation $x'_{i.} = (x'_{i1}, \ldots, x'_{i(TP)})$ ($i = 1, \ldots, N$), the criterion $\mathit{SSE_{kplusGeneralized}}$ can be computed as

@@

In general, the weighting scheme for k-plus anticlustering can become highly complex as it might encompass an arbitrary number of distribution moments, and hence, an arbitrary number of weights. While an exhaustive evaluation of the influence of weights is beyond the scope of the current paper, a small scale simulation study was conducted to investigate the influence of weights for a selected number of moments. The target was the optimization of $\mathit{SSE_{kplusGeneralized}}$ with $T = 4$, i.e. minimizing differences with regard to group means, variances, skewness and kurtosis. Two unweighted approaches were implemented by defining $\mathbf{w} = (1, 1, 1, 1)$. The unweighted standardization approach implemented a $z$-standardization of all columns---the original as well as the augmented variables---of the input data $\mathbf{X'}_{N \times \mathit{TP}}$ before the criterion $\mathit{SSE_{kplusGeneralized}}$ was optimized. In the second case, the unstandardized matrix $\mathbf{X'}_{N \times \mathit{TP}}$ was used. Four weighted approaches were implemented by giving a tenfold increased weight to one of the four moments, respectively. That is, to compute $\mathit{SSE_{kplusGeneralized}}$ favoring similarity in means during the optimization process, I set $\mathbf{w} = (10, 1, 1, 1)$; the weights favoring similarity in variances, skewness and kurtosis, were defined as $\mathbf{w} = (1, 10, 1, 1)$, $\mathbf{w} = (1, 1, 10, 1)$, and $\mathbf{w} = (1, 1, 1, 10)$, repectively. The weighted approaches did not use a standardization of the input data.

To compare the different weighting schemes, `r length(results)` data sets from were generated from a normal distribution ($M = 0$, $\mathit{SD} = 1$). The number of variables $P$ was determined randomly for each data set and varied between 1 and 5; the number of anticlusters $K$ was also determined randomly for each data set and varied between 2 and 4; the sample size $N$ randomly varied between 20 and 100 with the restriction that $N$ had to be divisible by $K$ so that equal-sized groups could be created. For each data set, $\mathit{SSE_{kplusGeneralized}}$ was optimized using the local maximum algorithm, once for each of the six competing weighting schemes.

```{r}
final_tab_sim1 <- round(Reduce("+", results) / length(results) * 100, 2)
rownames(final_tab_sim1) <- c(
  "$\\mathbf{w} = (1, 1, 1, 1)$ / unstandardized",
  "$\\mathbf{w} = (1, 1, 1, 1)$ / standardized",
  "$\\mathbf{w} = (10, 1, 1, 1)$",
  "$\\mathbf{w} = (1, 10, 1, 1)$",
  "$\\mathbf{w} = (1, 1, 10, 1)$",
  "$\\mathbf{w} = (1, 1, 1, 10)$"
)
colnames(final_tab_sim1) <- c(
  "$\\mathit{SSE}$",
  "$\\mathit{SSE_{Var}}$", 
  "$\\mathit{SSE_{Skew}}$",
  "$\\mathit{SSE_{Kurtosis}}$",
  "Average"
)
note <- paste0("Cells indicate how well the maximization $\\mathit{SSE_{kplusGeneralized}}$ fulfilled each of the four constituting criteria $\\mathit{SSE}$, $\\mathit{SSE_{Var}}$, $\\mathit{SSE_{Skew}}$, and $\\mathit{SSE_{Kurtosis}}$, depending on the weighting scheme. For each simulation run, the relative performance was computed as the $\\mathit{SSE}$ that was obtained by the given weighting scheme, divided by the best $\\mathit{SSE}$ that was obtained by any weighting scheme in this simulation run. Cell values represent the average relative performance across all ", length(results), " simulation runs.")

apa_table(
  final_tab_sim1, 
  escape = FALSE, 
  caption = "Relative performance of the competing weighting schemes in Simulation 1.",
  note = note,
  stub_indents = list("Unweighted approaches" = 1:2,
                      "Weighted approaches" = 3:6)
)

@@

Table 1 illustrates how well each of the six weighting schemes fulfilled each of the four constitutung criteria, averaged across the `r length(results)` simulation runs. First, the results show that for all weighting schemes, optimizing the combined criterion $\mathit{SSE_{kplusGeneralized}}$ yielded high performance with regard to all four criteria. Thus, the various k-plus criteria seem to align well when optimizing a combined criterion, which is good news for the k-plus approach to anticlustering. For the weighted approaches, the criterion that had received the highest weight was fulfilled to the highest degree, as should be expected. Interestingly, the unweighted standardization approach had the best average performance across the four criteria. While the degree of advantage over the other weighting schemes by itself does not seem to be remarkable, this pattern was observed in an astonishing `r mean(unstd_best) * 100`% of all simulation runs, while the level of chance would have been `r 1/6 * 100`% (`r format_p(sig_test$p.value)`). Among the competing weighting schemes, the unweighted standardization approach therefore yielded the most promising results. It also yielded satisfying results with regard to maximizing the standard $\mathit{SSE}$, comparable to the weighting scheme explicitly favoring $\mathit{SSE}$ (**TODO**: CHECK IF THESE RESULTS HOLD WHEN ALL DATA IS IN!). Hence, the unweighted standardization approach seemed to implicitly give high priority to minimizing differences in means. This is a pleasant result because, as discussed earlier, for the higher order moment $\mathit{SSE}$s to be valid, the group means have to be close to the overall mean, which is best accomplished if the standard $\mathit{SSE}$ has high priority.

@@

Two simulation studies and examples on real norming data showed that the k-plus criterion is well-suited to maximize between-group similarity with regard to multiple criteria. Given that similarity with regard to variances was achieved with almost no sacrifice to similarity in means, k-plus anticlustering should be considered as the default replacement of classical k-means when striving for between-group similarity.[^binarydata] The results did however show that fulfilling multiple criteria at the same time can be challenging in small data sets. In practice, users can try out different objectives and see which specification fulfills their needs. A reasonable approach is to start "greedy", e.g. by specifying that multiple distribution moments should equalized. If the results of such a greedy setup are not satisfying, users can reduce their requirements by opting for fewer criteria (e.g., only the mean and variance). Using the open source R package `anticlust`, the (repeated) application of k-plus anticlustering is easily accessible and free.

@@

The present paper focused on the new k-plus criterion for anticlustering, but only briefly considered the multicriterion nature of this criterion. A simple unweighted sum approach was used to optimize the k-plus objective, while more sophisticated algorithmic procedures exist to tackle multicriterion optimization problems [e.g., @brusco2009cross]. The sole focus on the objective function followed two reasonings. First, the applications and the simulations showed that the unweighted standardization approach yielded satisfying results that probably suffice in most applications. Moreover, it is possible that k-plus anticlustering has a fortunate solution space where many suitable partitions exist that fulfill multiple criteria to a satisfactory degree (as suggested by Figure 1). Still, more sophisticated approaches may yield theoretically improved results, as well as the opportunity to deliberately select among partitionings that may be preferred according to different k-plus objectives. @brusco2019 gave an example of how direct algorithms can be employed to optimize multiple criteria in anticlustering applications aiming for high within-group heterogeneity. Future research should investigate if k-plus anticlustering may profit from similar algorithmic treatment. A second reason for focusing on the k-plus objective per se---independently from the algorithm that is used to optimize it---was to obtain a fair comparison to existing anticlustering objectives. In the simulation study, differences in performance could not have been interpreted if the objective function as well as the algorithm used to optimize it had differed between anticlustering methods. It was of particular interest to compare the k-plus objective with the popular diversity objective. @papenberg2020 compared k-means anticlustering and diversity anticlustering[^clusterediting] in a simulation study and concluded that the diversity criterion should be preferred because k-means does not equalize the spread of the distribution between groups. In contrast, maximizing the diversity leads to an appropriate balance between location and spread. Using the k-plus extension, however, the k-means criterion has regained attractivity: When opting for similarity in means and variances---and for increasing $N$, skewness and kurtosis as well---, k-plus anticlustering clearly outperformed diversity anticlustering.

As a last outlook, it should be noted that because the covariance is defined as an expected value, the k-plus logic can be extended to minimize differences with regard to covariances, a possibility that was not investigated in the current paper. The covariance between two variables $x = (x_1, \ldots, x_N)$ and $y = (y_1, \ldots, y_N)$ is defined as $\frac{1}{N} \sum\limits_{i=1}^{N} (x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$. Thus, to maximize between-group similarity with regard to covariance structure, it is possible to augment the data set with an additional variable for each pair of variables $x$ and $y$, computed as $(x_i - \overline{\mathbf{x}}) (y_i -\overline{\mathbf{y}})$. Future research should examine how the k-plus approach can be adapted to including covariances, including a discussion of how to select weights for the possibly high number of additional variables. The unweighted standardization approach presented here may serve as a starting point for investigation.
